---
title: "Exercise answers"
author: ""
date: "December 3, 2015"
output: 
md_document:
variant: markdown_github
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, comment='')
```
  
1. Create a vector `simul_vec` of random numbers sampled from the distribution of your choice (Normal, Poisson, Binomial, etc).

    ```{r}
    simul_vec <- rnorm(500)
    ```

2. Create a function that returns a list with the mean and the standard deviation of the object `simul_vec`.

    ```{r}
    get_mean_sd <- function(x) {
      list(mean = mean(x), sd = sd(x))
    }
    ```

3. The function `rowMeans` returns the means of all rows in a data.frame. Create a function `sdMeans` that returns the sd of all rows (use the function `apply`). Read in the data [Mamiferos na Great Basin](http://ecologia.ib.usp.br/bie5782/lib/exe/fetch.php?media=dados:gbmam93.csv) and apply your function.

    ```{r}
    mammals <- read.csv("http://ecologia.ib.usp.br/bie5782/lib/exe/fetch.php?media=dados:gbmam93.csv")
    
    sdMeans <- function(df) {
      apply(df, 1, sd)
    }
    
    sdMeans(mammals)
    ```

4.  Read in the data [Levantamento em 3 caixetais do Estado de São Paulo](http://ecologia.ib.usp.br/bie5782/lib/exe/fetch.php?media=dados:caixeta.csv). With `ggplot2`, make a scatterplot of `h` vs. `cap`, faceting by `local`~`parcela` using the `facet_grid()` component. Try to use the argument `margins = TRUE` in `facet_grid()`. Add a `geom_smooth(method = lm)`.

    ```{r fig.width=8}
    library(ggplot2)
    
    caixetas <- read.csv("http://ecologia.ib.usp.br/bie5782/lib/exe/fetch.php?media=dados:caixeta.csv")
    
    ggplot(caixetas, aes(h, cap)) +
      geom_point() +
      geom_smooth(method = "lm") +
      facet_grid(local ~ parcela, margins = TRUE) +
      theme(axis.text.x = element_text(angle = 90))
    ```

5. Create a data.frame from the data `Titanic` (already loaded in your R session) using the function `as.data.frame()`. Group the data by different variables (or multiple variables at once) and take the average frequency of survivors in the grouped data (use the package `dplyr`).

    ```{r}
    library(dplyr)
    
    Titanic %>%
      as.data.frame() %>%
      group_by(Class, Age, Sex) %>%
      summarise(average_survived = mean(Freq))
    
    # proportion per Class
    library(tidyr)
    
    Titanic %>% 
      as.data.frame() %>%
      spread(Survived, Freq) %>%
      group_by(Class, Sex, Age) %>%
      mutate(prop_survived = Yes/sum(Yes, No)) %>%
      ungroup() %>%
      group_by(Class) %>%
      summarise(mean = mean(prop_survived, na.rm = TRUE))
    ```

6. Use the data below to create the data.frame `df2`

    ```{r}
    df2 <- read.table(text = "ind	sex	control	trt1	trt2
    1	m	7.9	12.3	10.7
    2	f	6.3	10.6	11.1
    3	f	9.5	13.1	13.8
    4	m	11.5	13.4	12.9", header = TRUE)
    ```

7. Convert `df2` to the long format with the function `gather` from the package `tidyr`. The columns `control`, `trt1` and `trt2` will be converted to 2 columns: 1) `treatment`, which will indicate the kind of condition in the experiment, and 2) `value` which will store the measurements for each observation.

    ```{r}
    df2 <- df2 %>% gather(treatment, value, control:trt2) 
    ```

8. Make a boxplot with `ggplot2` with the variable `sex` on the x axis, `value` on the y axis, and fill by `treatment`.

    ```{r}
    ggplot(df2, aes(sex, value, fill = treatment)) +
      geom_boxplot()
    ```