---
title: "version control with git and GitHub"
author: "Vitor Aguiar"
date: "December 4, 2015"
output: 
    md_document:
    variant: github_markdown
---

#### installing git:

- Mac OS (homebrew)
    - brew install git
    
- Ubuntu (apt-get)
    - apt-get install git
    
- Windows
    - [Windows build on the git website](https://git-scm.com)
    
#### settings

```
git config --global user.name "Your Name"
git config --global user.email "your@email.com"    
```

#### authentication with git remotes

consult the GitHub article on [adding SSH keys](https://help.github.com/articles/generating-ssh-keys/#platform-all)

#### create a repository on GitHub

- copy the URL (something like *git@github.com:username/repositoryname.git*)

#### in Rstudio:

- install the devtools package

```{r eval = FALSE}
install.packages("devtools")
```

- start a repository with:

```{r eval = FALSE}
devtools::create("PackageName")
```

Attention: R package names must start with a letter and only contain
alphanumeric characters, which means that underscores are not allowed.

#### go to your shell

- you need to start the git version control on the package directory:

```
cd path_to_package_dir
git init
git add remote origin https://github.com/user/repo.git
```

#### Rstudio

- go to the pane which lists your files
- click on the "Git" tab
- check the box under "Staged"
- click on the commit button
- write a message describing the commit
- click on "commit"
- back in the Rstudio pane, click on the "push" button

- go to the GitHub's repository and refresh the page

#### version control

- everytime you create or modify a file, you need to *stage* it
- *commits* represent snapshots of your project
- *push* sends commits to the remote repository
- *pull* retrives commits from the remote repository
